---
title: "ABA_Hackathon"
author: "Danya Zhang"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(magrittr)
library(lme4)
```

# Data Cleaning
```{r}
# read in the data
insurance <- read.csv("Medical Insurance dataset.csv")
```

```{r}
# check for null values
colSums(is.na(insurance)) # no nulls
```

```{r}
# new data set with altered variables ready for eda
insurance_clean <- insurance
insurance_clean$female <- ifelse(insurance$sex == "female", 1, 0) # binary var
insurance_clean$smoker <- ifelse(insurance$smoker == "yes", 1, 0) 
insurance_clean$region <- factor(insurance_clean$region)
insurance_clean %<>% 
  select(-sex)


insurance_clean <- insurance_clean %>%
  mutate(
  region = case_when(
  region == "northeast" ~ "NE",
  region == "northwest" ~ "NW",
  region == "southeast" ~ "SE",
  region == "southwest" ~ "SW",
    TRUE ~ region  
))

insurance_clean$region <- factor(insurance_clean$region)
```

```{r}
ggplot(insurance_clean, aes(x = age, y = charges)) + 
  geom_point() +
    labs(title = "Charges vs. Age", x = "Age", y = "Charges")

ggplot(insurance_clean, aes(x = bmi, y = charges)) + 
  geom_point() +
  labs(title = "Charges vs. BMI", x = "BMI", y = "Charges") +
  geom_smooth()
```

# Gamma Model
```{r}
# dependent variable is high right skewed so we try gamma
model_gamma <- glm(charges ~ ., data = insurance_clean, family = Gamma(link = "inverse"))
summary(model_gamma)
```

## Residual plot
```{r}
# Create a data frame with observed and predicted values
residuals_data <- data.frame(
  Observed = insurance_clean$charges,
  Predicted = fitted(model_gamma)
)

# Calculate residuals
residuals_data$Residuals <- residuals(model_gamma, type = "response")

# Create a residual plot
ggplot(residuals_data, aes(x = Predicted, y = Residuals)) +
  geom_point() +
  geom_smooth(method = "loess", formula = y ~ x, color = "red", se = FALSE) +
  labs(
    x = "Predicted Values",
    y = "Residuals",
    title = "Residual Plot for Gamma GLM"
  )
```


```{r}
# dependent variable is high right skewed so we try gamma
model_gamma2 <- glm(charges ~ female+age+bmi+children+smoker, data = insurance_clean, family = Gamma(link = "inverse"))
summary(model_gamma2)

anova(model_gamma2,model_gamma,test="LRT")

AIC(model_gamma2)
AIC(model_gamma)



# Create a data frame with observed and predicted values
residuals_data <- data.frame(
  Observed = insurance_clean$charges,
  Predicted = fitted(model_gamma2)
)

# Calculate residuals
residuals_data$Residuals <- residuals(model_gamma2, type = "response")

# Create a residual plot
ggplot(residuals_data, aes(x = Predicted, y = Residuals)) +
  geom_point() +
  geom_smooth(method = "loess", formula = y ~ x, color = "red", se = FALSE) +
  labs(
    x = "Predicted Values",
    y = "Residuals",
    title = "Residual Plot for Gamma GLM"
  )

# graph clearly shows a pattern, bad model, scrap
```

## Undersampling 
```{r}
# separate majority and minority classes 
majority_class <- filter(insurance_clean, smoker == 0)
minority_class <- filter(insurance_clean, smoker == 1)

set.seed(654)
majority_downsampled <- sample_n(majority_class, nrow(minority_class))
undersampled <- rbind(minority_class, majority_downsampled)
```

### Distribution of charges 
```{r}
undersampled %>%
  drop_na() %>%
  ggplot(aes(x = charges))+
  geom_density(alpha = 0.5) + 
  theme_minimal()+
  theme(title=element_text(size=10),strip.text = element_text(size=10),axis.title = element_text(size=7), plot.caption = element_text(size=9),axis.text.x=element_text(size=8))
```

### Gamma model on undersampled data 

```{r}
library(lme4)
library(glmm)
model_gamma_under <- glmer(charges ~ age + bmi + children + (1|smoker) + region + charges, data = undersampled, family = Gamma(link = "inverse"))
summary(model_gamma_under)
```

```{r}
# Create a data frame with observed and predicted values
residuals_data <- data.frame(
  Observed = undersampled$charges,
  Predicted = fitted(model_gamma_under)
)

# Calculate residuals
residuals_data$Residuals <- residuals(model_gamma_under, type = "response")

# Create a residual plot
ggplot(residuals_data, aes(x = Predicted, y = Residuals)) +
  geom_point() +
  geom_smooth(method = "loess", formula = y ~ x, color = "red", se = FALSE) +
  labs(
    x = "Predicted Values",
    y = "Residuals",
    title = "Residual Plot for Gamma GLM"
  )

# it's bad 
```


## Prediction
```{r}
# split data set into training, validation, and test
set.seed(654)
trainIndex <- createDataPartition(insurance_clean$charges, p = 0.7, 
                                  list = FALSE, 
                                  times = 1)
training_data <- insurance_clean[trainIndex, ]
leftover <- insurance_clean[-trainIndex, ]

# Now, split the remaining data into validation and test sets
validateIndex <- createDataPartition(leftover$charges, p = 0.5, 
                                    list = FALSE, 
                                    times = 1)
validation_data <- leftover[validateIndex, ]
test_data <- leftover[-validateIndex, ]
```

# Multilevel models
## varying intercepts constant slopes
```{r}
blah <- lmer(charges ~ age + bmi + children + region + female + (1 | smoker), 
     data = training_data)

residuals <- residuals(blah)

# Create a residual plot
ggplot(data = NULL, aes(x = fitted(blah), y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residual Plot") 
# varying intercepts constant slopes
coef(blah)
```


## transformed varying intercepts constant slopes
```{r}
blah <- lmer(sqrt(charges) ~ age + bmi + children + region + female + (1 | smoker), 
     data = training_data)

residuals <- residuals(blah)

# Create a residual plot
ggplot(data = NULL, aes(x = fitted(blah), y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residual Plot")
coef(blah)
```

## transformed varying intercept varying slopes
```{r}
blah <- lmer(sqrt(charges) ~ age + bmi + children + region + female + (age + bmi + children + region + female | smoker), data = training_data)


residuals <- residuals(blah)

# Create a residual plot
ggplot(data = NULL, aes(x = fitted(blah), y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residual Plot")

coef(blah)
```

## varying intercept varying slopes
```{r}
blah <- lmer(charges ~ age + bmi + children + region + female + (age + bmi + children + region + female | smoker), data = training_data)


residuals <- residuals(blah)

# Create a residual plot
ggplot(data = NULL, aes(x = fitted(blah), y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residual Plot")

coef(blah)
```

## constant intercept varying slopes
```{r}
blah <- lmer(charges ~ age + bmi + children + region + female + (age + bmi + children + region + female - 1 | smoker), data = training_data)

residuals <- residuals(blah)

# Create a residual plot
ggplot(data = NULL, aes(x = fitted(blah), y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residual Plot")

coef(blah)
```


## transformed constant intercept varying slopes
```{r}
blah <- lmer(sqrt(charges) ~ age + bmi + children + region + female + (age + bmi + children + region + female - 1 | smoker), data = training_data)

residuals <- residuals(blah)

# Create a residual plot
ggplot(data = NULL, aes(x = fitted(blah), y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residual Plot")

coef(blah)
```

## not bad
## varying slopes constant intercept
```{r}
blah <- lmer(sqrt(charges) ~  children + age:bmi + (region - 1 | smoker), data = training_data)
coef(blah)

residuals <- residuals(blah)

# Create a residual plot
ggplot(data = NULL, aes(x = fitted(blah), y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residual Plot")

coef(blah)

test_predictions_wog <- predict(blah, newdata = leftover)

test_rmse_wog <- sqrt(mean((test_predictions_wog - leftover$charges)^2))
cat("Test RMSE (without Gender):", test_rmse_wog, "\n")
```

## double levels
```{r}
blah <- lmer(charges ~ age + bmi + children + female + (1 | smoker) + (1 | region), data = insurance_clean)

residuals <- residuals(blah)

# Create a residual plot
ggplot(data = NULL, aes(x = fitted(blah), y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residual Plot")

coef(blah)
```

